---
title: "3. Kuhn Poker"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
library(RTutor)
RTutor::set.knit.print.opts(html.data.frame=TRUE, table.max.rows=30, round.digits=8, signif.digits=8)
library(gtree)
```

## Kuhn Poker Description

Joseph Kuhn (1950) developed a very simple poker game that is tractable for game theoretic analysis. Wikipedia describes the [Kuhn Poker game](https://en.wikipedia.org/wiki/Kuhn_poker) in conventional poker terms as follows:

+ Each player antes 1.

+ Each player is dealt one of the three cards, and the third is put aside unseen.

+ Player one can check or bet 1.
  + If player one checks then player two can check or bet 1.
    + If player two checks there is a showdown for the pot of 2 (i.e. the higher card wins 1 from the other player).
    + If player two bets then player one can fold or call.
      + If player one folds then player two takes the pot of 3 (i.e. winning 1 from player 1).
      + If player one calls there is a showdown for the pot of 4 (i.e. the higher card wins 2 from the other player).
  + If player one bets then player two can fold or call.
    + If player two folds then player one takes the pot of 3 (i.e. winning 1 from player 2).
    + If player two calls there is a showdown for the pot of 4 (i.e. the higher card wins 2 from the other player).

## Specifying as gtree game

We specify the game in gtree as follows:
```{r}
library(gtree)
game = org.game = new_game(
  gameId = "Kuhn-Poker",
  params = list(numPlayers=2),
  options = make_game_options(verbose=FALSE),
  stages = list(
    stage("dealCards",
      nature = list(
        # Player 1 gets a random card 1, 2, or 3
        natureMove("pl1_card", 1:3),
        # Draw from remaining cards for player 2
        natureMove("pl2_card", ~setdiff(1:3, pl1_card))
      )
    ),
    stage("pl1CheckBet",
      player=1,
      observe = "pl1_card",
      actions = list(
        action("cb1",c("check","bet"))
      )
    ),
    stage("pl2CheckBet",
      player=2,
      condition = ~ cb1 == "check",
      observe = c("pl2_card","cb1"),
      actions = list(
        action("cb2",c("check","bet"))
      )
    ),
    stage("pl2FoldCall",
      player=2,
      condition = ~ cb1 == "bet",
      observe = c("pl2_card","cb1"),
      actions = list(
        action("fc2",c("fold","call"))
      )
    ),
    stage("pl1FoldCall",
      player=1,
      condition = ~ is_true(cb1 == "check" & cb2=="bet"),
      observe = "cb2",
      actions = list(
        action("fc1",c("fold","call"))
      )
    ),
    stage("PayoffStage",
      player=1:2,
      compute=list(
        # Which player folds?
        folder ~ case_distinction(
          is_true(fc1 == "fold"),1,
          is_true(fc2 == "fold"),2,
          0 # 0 means no player folds
        ),
        
        # Which player wins?
        winner ~ case_distinction(
          folder == 1,2,
          folder == 2,1,
          folder == 0, (pl2_card > pl1_card) +1
        ),
        
        # How much gave each player to the pot?
        gave1 ~ 1 + 1*is_true((cb1 == "bet") | (fc1 == "call")),
        gave2 ~ 1 + 1*is_true((cb2 == "bet") | (fc2 == "call")),
        pot ~ gave1 + gave2,
        
        # Final payoffs
        payoff_1 ~ (winner == 1)*pot - gave1,
        payoff_2 ~ (winner == 2)*pot - gave2
      )
    )
  )
) 
```

To better understand the definition and to check whether we have correctly specified the game, it is useful to take a look at the outcomes:
```{r table.max.rows = 6}
game %>% get_outcomes() %>% head(6)
```
Look at the first row. We see from `cb1` and `cb2` that this corresponds to an outcome in which both players check. The variables `fc2` and `fc1` take `NA` values because there is no decision to fold or call if both players check.

Formulas in the game definition will be internaly evaluated in a vectorized fashion over similar data frames and may take `NA` values. The helper function `is_true` takes a logical vector and replaces `NA` values with `FALSE`. I use this function in the game definition where a condition must evaluate to either `TRUE` or `FALSE` while `NA` values are not allowed.

You may also take a look at the definition of `pl2_card` in the first stage. Here the `set` of the random variable is a formula and depends on the previously computed value of `pl1_card`.


Let us also take a look at the game size:
```{r}
game %>%
  game_print_size_info()
```
While the number of pure strategy profiles is not really small, the game still seems of tractable size for numerical analysis.

## Solving Kuhn Poker

Let us now solve the game using the `gambit-logit` solver, which is the default solver for finding a mixed strategy equilibrium:
```{r}
game %>%
  game_gambit_solve(mixed=TRUE)
```

Let us first take a look at the expected equilibrium outcomes:
```{r}
game %>% 
  eq_expected_outcomes() %>% 
  select(payoff_1,payoff_2, cb1, fc1, cb2,fc2)
```

```{r}
game %>%
  eq_cond_expected_outcomes(pl1_card = 3, cb1=c("bet", "check")) %>% 
  select(payoff_1,payoff_2, cb1, fc1, cb2,fc2)

```

```{r}
game %>%
  eq_cond_expected_outcomes(pl1_card = 1, cb1=c("bet", "check")) %>% 
  select(payoff_1,payoff_2, cb1, fc1, cb2,fc2)
```



For more details let us look at the equilibrium tables:
```{r}
game_print_eq_tables(game)
```

```{r}
fgame = org.game %>%
  game_copy() %>%
  game_fix_actions(cb1="check", tremble.prob = NULL) %>%
  game_print() 

fgame %>% 
  game_gambit_solve(mixed=TRUE) %>%
  eq_expected_outcomes() %>%
  select(payoff_1,payoff_2, cb1, fc1, cb2,fc2)  


fgame %>% 
  game_gambit_solve("gambit-logit -e") %>%
  eq_expected_outcomes() %>%
  select(payoff_1,payoff_2, cb1, fc1, cb2,fc2)  

fgame %>%
  eq_cond_expected_outcomes(pl1_card = rep(1:3, each=2), cb1=rep(c("bet", "check"), times=3)) %>% 
  select(pl1_card, payoff_1,payoff_2, cb1, fc1, cb2,fc2)


pref = pref_custom(util_1 = ifelse(cb1=="check", payoff_1+1000, payoff_1), util_2 = payoff_2)
game %>%
  game_set_preferences(pref) %>%
  game_gambit_solve(mixed=TRUE) %>%
  eq_expected_outcomes()  %>%
  select(payoff_1,payoff_2, cb1, fc1, cb2,fc2)  
  

  
```


##  Loss Averse Players
What if players are averse?
```{r}
game %>%
  game_set_preferences(pref_lossAv(lambda=3, r=0)) %>%
  game_gambit_solve(mixed=TRUE) %>%
  eq_expected_outcomes() %>% 
  select(payoff_1,payoff_2, cb1, fc1, cb2,fc2)
```


## References

Kuhn, H. W. (1950). "Simplified Two-Person Poker". In Kuhn, H. W.; Tucker, A. W. (eds.). Contributions to the Theory of Games. 1. Princeton University Press. pp. 97–103.
