---
title: "4. Comparing with Gambit Python API: QRE in a Sender-Receriver Game"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
library(gtree)
```

## A Sender-Receiver Game: Gambit's Python API vs gtree

[Here is a jupyter notebook](https://nbviewer.jupyter.org/github/gambitproject/gambit/blob/master/contrib/samples/sendrecv.ipynb) by Valeria Burdea that creates a sender-receive game using [Gambit's Python interface](http://www.gambit-project.org/gambit16/16.0.0/pyapi.html). For a comparison, this tutorial shows how one can create and analyse the same game in a relatively simple fashion using `gtree`.

I just copy the game description from the Jupyter notebook:

"This is a 2-player sequential game the structure of which is inspired by Glazer and Rubinstein (2004, 2006). In the first stage, the sender is dealt 2 cards - one card orange, one card blue. Each card can take a value between 1 and 9; each combination of values is equally likely. A hand is good if the sum of the two cards is at least 9; a hand is bad otherwise. Upon observing the two cards, the sender chooses one to reveal to the receiver. The receiver observes the revealed card and chooses an action between 'Accept' and 'Reject'. The incentives are such that the sender would always want the receiver to accept, while the receiver would only want to accept if the hand is good and reject otherwise."

Here is the implementation of the game using `gtree` (the exact payoffs were taken from the Python code):
```{r}
library(gtree)
game = new_game(
  gameId = "Sender-Receiver",
  params = list(numPlayers=2, maxVal=9),
  options = make_game_options(verbose=FALSE),
  stages = list(
    stage("DrawStage",
      nature = list(
        natureMove("orange", ~1:maxVal),
        natureMove("blue", ~1:maxVal)
      )
    ),
    stage("RevealStage",
      player=1,
      observe = c("orange","blue"),
      actions = list(
        action("reveal",c("orange","blue"))
      )
    ),
    stage("AcceptStage",
      player=2,
      compute = list(
        shownVal ~ ifelse(reveal=="orange", orange, blue)
      ),
      observe = c("shownVal"),
      actions = list(
        action("accept",c(TRUE, FALSE))
      )
    ),
    stage("PayoffStage",
      player=1:2,
      compute=list(
        hand ~ ifelse(orange+blue>=maxVal, "good","bad"), 
        payoff_1 ~ ifelse(accept, 1,0), # Sender wants accept
        payoff_2 ~ case_distinction(
          accept & hand=="good", 1,
          accept & hand=="bad", 0,
          !accept & hand=="good",0,
          !accept & hand=="bad", 1
        )
      )
    )
  )
) 
```

I would argue that gtree's game definition using stages follows in a straightforward and simple fashion the verbal game description. 

Gambit's Python API is instead designed to directly specify game trees. For comparison, take a look [here](https://nbviewer.jupyter.org/github/gambitproject/gambit/blob/master/contrib/samples/sendrecv.ipynb).

## Huge number of strategy profiles

Shall we try to compute all pure SPE?
```{r}
game %>% game_solve()
```
Ups... looks like the game is too large to be solved. Here is some additional size information.
```{r}
# Get some size information about the game
game %>% 
  game_print_size_info()
```
The game has a *gigantic* number of strategy profiles. 

## Only one subgame???

Interestingly, the game also has only 1 subgame, even though player 1 perfectly observes the values of the orange and blue card. For an illustration, why there is only one subgame, we export a smaller version of the game in which cards only take values of 1 or 2 to a Gambit efg file.
```{r}
game %>%
  game_change_param(maxVal = 2) %>%
  game_write_efg("sender_receiver.efg")
```
I have manually opened the efg file with Gambit and exported the following game tree as an svg graphic:
![sender receiver gametree](sender_receiver.svg)

Recall the definition of a *subgame*, e.g. from [Wikipedia](https://en.wikipedia.org/wiki/Subgame):

1. It has a single initial node that is the only member of that node's information set (i.e. the initial node is in a singleton information set).
2. If a node is contained in the subgame then so are all of its successors.
3. If a node in a particular information set is in the subgame then all members of that information set belong to the subgame.

Even though the moves of player 1 start at singleton information sets (Condition 1), candidate subgames starting at these nodes violate Condition 3.

## Using the gambit-logit solver from gtree

If we cannot reduce the number of relevant strategy profiles by subgames, our internal gtree solver cannot practically solve games with so many strategy profiles. However, Gambit has solvers that can find at least one equilibrium of this game. Here we let `gambit-logit` do its magic (you can fetch a cup of coffee until its finished):
```{r}
game %>%
  game_change_param(maxVal = 4) %>%
  game_print_size_info() %>%
  game_gambit_solve(mixed=TRUE,gambit.command = "gambit-logit -q -e")

game %>%  eq_tables()
```
The results seem intutive: Player 2 only accepts if and only if the shown card is at least a 6. Player 1 always shows a card that is at least 6 if he has one.

