---
title: "gtree - Exploring Reciprocity Preferences"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
library(gtree)
```

## An Ultimatum Game with possibility of random offer

```{r}
library(gtree)
game = new_game(
  gameId = "UG_RandomOffer",
  options = make_game_options(verbose=TRUE),
  params = list(numPlayers=2,natureProb = 0.5),
  stages = list(
    stage("whoOffers",
      nature = list(
        natureMove("offerType",c("nature","player"), probs = ~c(natureProb, 1-natureProb))
      )  
    ),
    stage("randomOffer",
      condition = ~ offerType == "nature"
      nature = list(
        natureMove("natureOffer",0:1)
      )
    ),
    stage("proposerStage",
      player=1,
      condition = ~ offerType == "player",
      actions = list(
        action("playerOffer",0:1)
      )
    ),
    stage("responderStage",
      player=2,
      observe = "offer",
      actions = list(
        action("accept",c(FALSE,TRUE))
      )
    ),
    stage("PayoffStage",
      player=1:2,
      compute=list(
        payoff_1 ~ ifelse(accept, cake-offer,0),
        payoff_2 ~ ifelse(accept, offer,0)
      )
    )
  )
)
```


The `gameId` should be a unique name of the game. This is relevant if we want to conveniently save results, like computed equilibria, in the default folder structure used by `gtree`.

We then define a list of parameters `params`. Parameters can be referenced to in later definitions of the game. Note that you must *always* specify the number of players in a parameter called `numPlayers`. We also specify the size of the `cake` that can be distributed between the proposer and responder in the game.

Heart of our definition is a list of 3 `stages`. Each stage in the list should be generated with the function `stage` that sets defaults and transforms all formulas into a canoncial format.

1. In the first stage is named `proposerStage`. The argument `player=1`, specifies that player 1 acts here.
She chooses an action `offer`, that is created with the function `action` and element of a list `actions`. The function `action` first requires a name and then a set of possible values the action can take. Here we specify the set as a formula `~ 0:cake`. This means we compute the action set based on the specified parameters and possibly based on previously computed variables including chosen action values or realized moves of nature.<br>
Alternatively, we could also provide a fixed action set without formula e.q. `0:4`. This can not contain references to parameters or variables of the game and is always fixed when the game is created.

2. In the second stage player 2, observes the offer. This is specified by the argument `observe="offer"`. The argument observe specifies all observed variables as a simple character vector, or remains `NULL` if nothing is observed.<br>
Player 2 then decides whether to accept the action or not. Here we have chosen the fixed action set `c(FALSE,TRUE)`. You could encode specify the set for accept in a different way, e.g. as a character vector `c("reject","accept")` or an integer vector `c(0,1)`. 


3. The third stage just computes variables as specified by the list provided for the field `compute`. You can briefly specify a computation with the formula syntax `name ~ formula`.  Note that for each player `i` you must compute somewhere in your game the variable `payoff_i`, like `payoff_1` and `payoff_2`, that specifies the (monetary) payoff for that player. We can later easily transform these monetary payoffs, using some alternative outcome based utility function, e.g. to account for inequality aversion or loss aversion.
<br>You can use any vectorized, deterministic R function to specify a computed variable. Random variables must be declared separately, as a move of nature, however (see further below). Here we use the function `ifelse` for a simple distinction of two cases. For distinguishing more than two cases the functions `cases` in `gtreeCore` provides a simple syntax. Note that we could have more compactly written:
`payoff_1 ~ (cake-offer)*accept` and `payoff_2 ~ offer*accept`
<br>For computing equilibria, it does not really matter which players you specify a stage in which no action takes place. However, `gtree` also has (currently rudimentary) features to run a game as a web-based experiment. When running as an experiment, a stage will be shown to all players that are specified in the `players` field. If an action is taken in a stage, exactly ONE player must be specified. For stages without actions, you can specify any number of players including no player.

We can get a short overview of a specified game by typing its variable name in the R console.
```{r}
game
```

## Compiling and Solving for equilibria

In order to compute equlibria `gtree` will create different internal representations of the game. While the function `game_solve_spe` will automatically call the corresponding functions, it is useful to call them manually before.
```{r}
game_compile(game, for.internal.solver = TRUE)
game
```
We now see some additional information about the size of the game in terms of number of outcomes, information sets, subgames and number of pure strategy profiles.

As a game tree our game looks as follows:
![ug.svg](ug.svg)


Remark 1: By default `game_compile` only computes the information neccessary to create a game tree that can be saved as a Gambit `.efg` file and then solved via Gambit. For finding pure strategy equilibria, usually (but not always) the internal `gtree` solver is faster. It will compute some additional information, e.g. identifying in which information sets new subgames start. The argument `for.internal.solver` forced the computation of this additional information.

Remark 2: To generate an image of the game tree, we can export the game to a [Gambit](http://www.gambit-project.org/) extensive form game format using the following command:
```{r eval=FALSE}
game_write_efg(game,"UG.efg")
```
We can then open the file with Gambit GUI, which draws the game tree.

Ok, enough remarks. Let us now solve the game. More precisely, we use the internal solver to find all pure strategy subgame (SPE) perfect equilibria.
```{r}
game_solve_spe(game)
```

### Equilibrium outcomes

The following code shows the equilibrium outcomes, i.e. all actions and computed variables on the equilibrium path. 
```{r}
game.eq.outcomes(game)
```
We have two different equilibrium outcomes: the proposer either offers 
0 or 1 and in both equilibrium outcomes the offer will be accepted. For games with moves of nature there is also a function `game.expected.eq.outcomes` that shows expected equilibrium outcomes.

Remark on my naming convention for functions: A function that starts with `game.` takes as first argument a game object but does not return a game object. In contrast, a function starting with `game_` takes a game object as first argument and always returns a game object.


### Equilibria represented as action tables

An equilibrium also describes equilibrium play off the equilibrium path, e.g. it also describes whether player 2 would `accept` an out-off-equilibrium `offer` of 3.

In gtree there are different ways to represent the computed equilibria. Here is a convenient representation for pure strategy equilibria:
```{r}
game.eq.tables(game, combine=2, reduce.tables = TRUE)
```
We have a list with a `tibble` for every action variable.

  - The first table describes the equilibrium offers: In the first equilibrium the offer is 1 and in the 2nd it is 0.
  
  - The second table describes the conditional `accept` decisions: In the first equilibrium an offer of `0` is rejected, in all other cases the offer is accepted.
  
The argument `combine` can take the values 0,1 and 2 and describes how the results of different equilibria are combined. For example, with `combine = 0`, we would get a separate list for every equilibrium. The argument `reduce.tables` automatically removes key columns that have no impact on the chosen equilibrium action.

In my view that such a representation of equilibria is quite intuitive and convenient when comparing equilibrium predictions with experimental results.

### Internal representation of equilibria

Yet, `gtree` uses  a different internal representation of equilibria. To understand the internal representation it is first useful to show all possible outcomes of the game:
```{r}
game.outcomes(game, reduce.cols=TRUE)
```
This data frame corresponds to all possible full paths that can be taken through the game tree. (It does not fully describe the game tree, though, since it contains no specification of information sets.)

Let us now show the internal representation of our 2 equilibria:
```{r}
game.eq.li(game)
```
It is a list with a matrix for each equilibrium. Each row corresponds to one possible outcome of the game and the column describe for each action the equilibrium choice probability on the corresponding outcome path. The last column specifies the total probality of the particular outcome in the equilibrium. 


The attribute `info.set.probs` shows the most compact equilibrium representation. It is just a numerical vector that describes the move probability for every possible move in every information set. This is similar to the equilibrium representation that you get if you manually call a Gambit solver on an `.efg` file (except that Gambit has a different default ordering of the information sets). The information sets are further described in the `game` object. In principle you can access the information, e.g. by typing
```{r}
game$tg$ise.df
```
Yet, there should not be any need to dig so deeply into the internal game representation of `gtree`.

## Setting different preferences

So far we assumed that the specified payoffs `payoff_1` and `payoff_2` are equal to players' utility. One motivation for gtree is to conveniently solve games for different specifications of players' preferences that can account e.g. for inequality aversion or loss aversion.

While in principle, one could account for different outcome based preferences by directly adapting the formulas for `payoff_1` and `payoff_2` in the game definition, we prefer a slightly different approach.

In the preferred approach the specified payoffs in the game definition are interpreted as monetary or material payoffs. This means games created by `new_game` can very closely match the structure of economic experiments, for which we only know the specified monetary payoffs. 

After the game is specified, we can use the function `game_set_preferences` to specify a utility function for which we want to find equilibria.

For example, consider the following inequality aversion utility function (Fehr and Schmidt, 1999)
\[
u_i = \pi_i 
  - \alpha \frac {1}{n-1}\sum_{j \ne i} \max(\pi_j - \pi_i,0)
  + \beta \frac {1}{n-1}\sum_{j \ne i} \max(\pi_i - \pi_j,0)
\]
where $\pi$ denotes monetary payoffs.

The following code manually specifies these preferences and solves for subgame perfect equilibria:
```{r}
prefs = pref_custom(
  util_1 = payoff_1 - alpha*(pmax(payoff_2-payoff_1,0)) - beta*(pmax(payoff_1-payoff_2,0)),
  util_2 = payoff_2 - alpha*(pmax(payoff_1-payoff_2,0)) - beta*(pmax(payoff_1-payoff_2,0)),
  params = list(alpha = 1, beta=0.5),
  label = "ineqAv_100_50"
)

game %>%
  game_set_preferences(prefs) %>%
  game_solve_spe(game) %>% 
  game.eq.tables()
```
We see that with inequality aversion with an envy parameter of `alpha=1` and a guilt parameter of `beta=0.5` there is a unique SPE in which the proposer offers half of the cake.

Some common preference clases that are only transformations of material payoffs are included into gtree. All functions start with the prefix `pref_`. The following code verifies that guilt is not essential for positive offers by the proposer.

```{r}
game %>%
  game_set_preferences(pref_ineqAv(alpha=1,beta=0, player=1:2)) %>%
  game_solve_spe(game) %>% 
  game.eq.tables()
```

We will discuss later how one can specify heterogenous preferences via different preference types.

## Moves of Nature

The following code generates a game that has an initial stage with a move of nature, a.k.a. a random variable, `autoAccept`. With probability 20% an offer will always be automatically accepted, even if the responder wants to reject it. This is like a probabilistic mixture between an Ultimatum Game and an Dictator Game.

The effect of `autoAccept` is  specified in the `payoffStage`. No player observes `autoAccept`.
```{r}
game = new_game(
  gameId = "UG_AutoAccept",
  params = list(numPlayers=2, cake=4),
  options = make_game_options(verbose=FALSE),
  stages = list(
    stage("autoAcceptStage",
      nature=list(
        natureMove("autoAccept", c(FALSE,TRUE),probs = c(0.8,0.2))
      )
    ),
    stage("proposerStage",
      player=1,
      actions = list(
        action("offer",~0:cake)
      )
    ),
    stage("responderStage",
      player=2,
      observe = "offer",
      actions = list(
        action("accept",c(FALSE,TRUE))
      )
    ),
    stage("PayoffStage",
      player=1:2,
      compute=list(
        payoff_1 ~ ifelse(accept | autoAccept, cake-offer,0),
        payoff_2 ~ ifelse(accept | autoAccept, offer,0)
      )
    )
  )
) 

game %>%
  game_solve_spe() %>%
  game.eq.tables()

```

If you are not familiar with the exact definition of a subgame perfect equilibrium (and maybe even if you are), the result may surprise you. We have 32 different pure SPE and even though players care only about their own monetary payoff, every offer between 0 and 3 is supported by some SPE. 

We also see that every possible offer between 0 and 4 is accepted in half of the SPE and rejected in the other half.

Let us take a look at the game structure:
```{r}
# Show size information only 
print(game, show.stages=FALSE, show.eq=FALSE)
```

We only have a single subgame. The reason is that a new subgame only begins if a player perfectly knows the outcomes of all previous moves of natures and all previous actions. Since no player observes the initial realization of `autoAccept` the only "subgame", is the complete game. This means the set of SPE is equal to the set of Nash equilibria and backward induction does not help to compute SPE. You may recall that the ultimatum game has a lot of Nash equilibria supported by out-off-equilibrium threats to reject out-off-equilibrium offers.

(As a test question you can ask yourself why in our game there is no SPE in which the proposer offers the whole cake, even though that is a NE in the original ultimatum game.)

Intuitively, most of us would still solve the game via some form of backward induction. Formally, one could choose a refinement like `sequential equilibrium` or `trembling-hand perfect equilibrium` in which only equilibrium offers of 0 or 1 would survive. Unfortunately, I don't know a fast algorithm to compute sequential, trembling-hand perfect or perfect Bayesian algorithms (neither is there a Gambit command line tool for it).

We can however modify the game to manually add trembles in order to select equilibria that will usually coincide with trembling-hand perfect equilbria. The call to function `game_add_tremble` in the following code changes the game structure such that with probability 0.01%, the proposer trembles and picks a random offer.
```{r}
game %>%
  # make deep copy since game structure will be changed  
  game_copy() %>%  
  game_add_tremble(action="offer", tremble.prob = 0.0001) %>%
  game_solve_spe() %>%
  print()
```
You see that 2 new stages have been automatically added before the `proposerStage`: `proposeStage_willTremble` determines whether the proposer will tremble and `proposerStage_performTremble` will choose a random offer in case trembling happens.

The stages `proposerStage` and `proposeStage_performTremble` have a `condition` field that specifies whether the stage will be played or not. Here it simply depends on whether trembling happens or not.

The number of possible outcomes naturally increases, but the number of strategy profiles has not increased by the possibility of trembles.

Looking at the equilibria, we see that only the two intuitive equilibria where the proposer offers 0 or 1 survive.  

## Optimal position of moves of nature in stage game

Instead of adding trembles there would be a much simpler modification to reduce the number of equilibria that follows a simple rule:

    Define moves of nature as late as possible in the game.

Since the variable `autoAccept` is first and only used in the last stage, we should not specify before that last stage. So consider the following modification of our game. 

```{r}
game = new_game(
  gameId = "UG_AutoAccept",
  params = list(numPlayers=2, cake=4, acceptProb=0.2),
  options = make_game_options(verbose=FALSE),
  stages = list(
    stage("proposerStage",
      player=1,
      actions = list(
        action("offer",~0:cake)
      )
    ),
    stage("responderStage",
      player=2,
      observe = "offer",
      actions = list(
        action("accept",c(FALSE,TRUE))
      )
    ),
    stage("PayoffStage",
      player=1:2,
      nature=list(
        natureMove("autoAccept", c(FALSE,TRUE),probs = ~c(1-acceptProb,acceptProb))
      ),
      compute=list(
        payoff_1 ~ ifelse(accept | autoAccept, cake-offer,0),
        payoff_2 ~ ifelse(accept | autoAccept, offer,0)
      )
    )
  )
) 

game %>%
  game_solve_spe() %>%
  print(show.stages = FALSE)
```

We now have again 6 subgames and subgame perfection has bite again. Only the two equilibria where 0 or 1 is offered are selected.

Note that in a stage first move of natures are computed, then deterministic variables are computated and then actions take place. This means the payoffs in the `PayoffStage` can use the value of `autoAccept`. Otherwise, we simply had to add a separate stage.
